\documentclass[11pt]{article}
\usepackage[top=1.00in, bottom=1.0in, left=1.1in, right=1.1in]{geometry}
\renewcommand{\baselinestretch}{1.1}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{parskip}

\def\labelitemi{--}
\parindent=0pt


\title{Fuelinex Climate Chambers notes}
\author{Christophe Rouleau-Desrochers }
\date{\today}

\begin{document}
\maketitle	
\pagebreak

% Optional TOC
% \tableofcontents
% \pagebreak

%--Paper--

\section{Meeting with Patrick (May 30, 2024)}
\subsection {Burnt leaves and schedule improvements}
\textbf{Context}: during the CoolSpring treatment, the trees were in the chambers for several weeks. The temperature in the day was at a maximum of 8.2C. I changed the light schedule to the sunrise and sunset of that week in Vancouver. However, I didn't ramp the light intensity. In April, as most of the trees have leafed out, some of them showed yellow leaves, not because they were young. It looked more like chlorosis. I asked Patrick if that could be because the light intensity was too high.
\par \textbf {Question}: how could we prevent the leaves from turning yellow in the future? 
\par textbf  {Answer} : I should change the schedule so the light intensity is ramped. I could either make many more entries (which makes things more complicated) or just ramp in the option section. Because the morning and night lights are not as strong as during the middle of the day, this would represent the outside light better.
\subsection {Drainage problem}
\textbf{Context}: last summer (2023), we had problems with the drainage. None or almost no water would come out from the pipes. The heat wave treatment especially was a problem where the relative humidity was a 100\%. Also, during the spring treatment, one of the medium chamber was smelling bad.
\textbf{Question} : how can we get the drain from working again?
\textbf{Answer}: Unclug the drains by removing the floor metal section. Also, clean in air intake filter. To do this, I need to remove the plastic case and clean the filter. They have youtube videos on how to maintain the chambers.


\section{Meeting with Roman (June 24, 2024}
\textbg {Things done}
\begin {itemize}
	\item Fixed the not far red lights that were not working in the walkin chambers, as well as in the smaller chambers
	\item He showed me how to clean the filters. 
	\item He fixed the electric plug that wasn't working in the walk in chamber
	\item We saw that the drains were clugged, I'll need to unglug them.
	\item We noticed that the copper pipes weld. I think it's not worth fixing them, as the room drain is close enough from the chambers
\end {itemize}
%--/Paper--

\end{document}